\textbf{\supplement}\vskip 2ex \mytitle\vskip 2ex

### 2022-01-27 paper-seb-main.tex
Fixing final issues for PLOS

\input generated-info-for-main.tex
%\IfFileExists{generated-info-for-main.tex}{\input{generated-info-for-main.tex}}{\typeout{No generated-info-for-main.tex - error ignored, but you need to latex paper-seb-supplementary-material.tex to generate it}}

@@

In short, computers, data, and computer code are central to modern science. However, using code raises many critical questions: formats, backup, cyber-vulnerability, version control, integrity checking (e.g., managing human error), auditing, debugging and testing, and more. All are non-trivial concerns requiring expertise to do reliably. Results claimed, then, however nicely they may be expressed, are no more reliable than the code and data that was used to generate them. A common oversight in scientific papers is to present a mathematical model, such as a set of differential equations, but omit how that model is transformed into code that generates the results the paper summarizes.

@@

Just as in statistics, code and results from code need to be discussed and presented in a way that ground belief in claims derived from using them. Specifically, code should be developed in a sufficiently professional and rigorous way that is able to support clear presentation and scrutiny. Developing justifiably reliable code is the concern of the field of \emph{software engineering}, which will be discussed further below in this paper, as well as more substantially in this paper's \supplement).

To make critical claims, to inform public policy or to support further research, models need to be run under varying assumptions \cite{whitty}. Being able to understand (at least in principle) the exact code used in implementing a model is critical to having confidence in the results that rely on it. Unfortunately code is rarely considered a substantial part of the science to which it contributes.

We would not believe a statistical claim that was obtained from some \emph{ad hoc\/} analysis with a new fangled method devised just for one project --- instead, we demand statistics that is recognizable, even traditional, so we are assured we understand what has been done and how reliable results were obtained.

This paper will show that unreliable computational dependencies in science are widespread and serious. Furthermore, code is rarely published in any useful form or professionally scrutinized, and therefore the code itself not contribute to furthering science, for instance through replication or reproduction.

In short, the quality of much modern science is undermined because the code it relies on is rarely of adequate quality \emph{for the uses to which it is put}. This paper will explore the extent of these problems and reasons why they persist. The paper additionally suggests some ways forward. The \supplement\ is an integral part of the suggested solutions. In particular, the \supplement\ section~\ref{supplementary-Speigelhalter-section} summarizes Speigelhater's uncontroversial statistics probes and draws explicit analogies for the critical assessment of the quality of scientific code.

@@

{\raggedright
\initialiseBibliography{References}{1}{}
\bibliographystyle{plos2015}
\bibliography{paper-seb-main.bib}

>>>>

@@

In short, computers, data, and computer code are central to modern science. However, using code raises many critical questions: formats, backup, cyber-vulnerability, version control, integrity checking (e.g., managing human error), auditing, debugging and testing, and more. Software code, like statistics, is also subject to unintentional bias \cite{se-bias}.  All these issues are non-trivial concerns requiring technical expertise to manage well.

A common oversight in scientific papers is to present a model, such as a set of differential equations, but omit how that model is transformed into code that generates the results the paper summarizes; if so, the code may have problems that cannot be identified as there is no specification to reference it to.

@@

We would not believe a statistical claim that was obtained from some \emph{ad hoc\/} analysis with a new fangled method devised just for one project --- instead, we demand statistics that is recognizable, even traditional, so we are assured we understand what has been done and how reliable results were obtained. An interesting overlap with statistical and software engineering scrutiny is the many papers that just disclose as part of their methodology that they used a particular statistical package (e.g., ``Data analyses were performed using SAS 9.2 (SAS Institute, Cary, North Carolina, USA)\@.'' yet \emph{how\/} those analyses might have been performed is not discussed. The problem is that admitting using SAS 9.2 or any other named system does not really help scrutiny, as such systems can do anything with the data. A reviewer, if nobody else, needs to actually examine the statistical code itself and its documentation to assess whether the analysis presented in the paper is reliable.

It is recognized that to make critical claims, models need to be run under varying assumptions \cite{whitty}, yet somehow it is easy to overlook that the code that implements the models also needs to be carefully tested under varying assumptions to uncover and fix bugs and biases. Being able to understand (at least in principle) the exact code used in implementing a model is critical to having confidence in the results that rely on it. Unfortunately code is rarely considered a substantial part of the science to which it contributes.

This paper reviews a selection of papers in leading international journals, and finds that both papers and journal policies take code for granted. This paper then argues that, just as is routine for statistics, code and results from code (and the data it is run on) need to be discussed and presented in a way that properly assure belief in any claims derived from using them. Specifically, code should be developed and discussed in a sufficiently professional, rigorous, and recognizable way that is able to support clear presentation and scrutiny. Developing justifiably reliable code is the concern of the field of \emph{software engineering}, which will be discussed further below, as well as more substantially in this paper's \supplement).

This paper shows that unreliable computational dependencies in science are widespread. Furthermore, code is rarely published in any useful form or professionally scrutinized, and therefore the code itself does not contribute to furthering reliable science, for instance through replication or reproduction. In short, the quality of much modern science seems to be undermined because the code it relies on is rarely of adequate quality \emph{for the uses to which it is put}.

This paper explores the extent of these software engineering problems in published science. The paper additionally suggests some ways forward. The \supplement\ is an integral part of the suggested solutions. In particular, the \supplement\ section~\ref{supplementary-Speigelhalter-section} summarizes Speigelhater's uncontroversial statistics probes and draws explicit analogies for the critical assessment of the quality of scientific code.

@@

{\raggedright
\initialiseBibliography{References}{1}{}
% \bibliographystyle{plos2015} - already provided in plos-header.tex
\bibliography{paper-seb-main.bib}

### 2022-01-27 paper-seb-supplementary-material.tex
Fixing final issues for PLOS

\makeatletter
      \renewcommand{\thefootnote}{\arabic{footnote}}
      {\let\oldbibcite=\bibcite
      % \bibcite{bad-code}{{31}{2020}{{Richards and Boudnik}}{{Richards and Boudnik}}}
      \newcount\mainNumberOfReferences \mainNumberOfReferences=0
      \def\bibcite#1#2{\global\advance\mainNumberOfReferences by 1\oldbibcite{#1}{#2}}
      \input paper-seb-main.aux
      %\inputifexists{paper-seb-main.aux}{You need to run Latex on paper-seb-main.tex because we need its aux file}
      }
\makeatother

@@

\noindent\vbox{% The table of contents
\def\contents#1#2{\vskip 0mm \noindent{\bfseries #1} \dotfill\ #2\\}
\def\subcontents#1#2{\hbox to 1em{}{#1} \dotfill\ #2\\}
\noindent\input generated-supplementary.toc
%\inputifexists{generated-supplementary.toc}{you need to Latex supplementary-material.tex again}
}
\newwrite\contentsfile
\immediate\openout\contentsfile=generated-supplementary.toc

@@

\vbox{\noindent\input generated-minicontents.tex
%\inputifexists{generated-minicontents.tex}{You need to run Latex again on supplementary-material.tex to generate the minicontents file}
}

@@

\immediate\closeout\infofile
\immediate\closeout\contentsfile
\immediate\closeout\minicontentsfile
\end{document}

>>>>

\makeatletter
      \renewcommand{\thefootnote}{\arabic{footnote}}
      {\let\oldbibcite=\bibcite
      % \bibcite{bad-code}{{31}{2020}{{Richards and Boudnik}}{{Richards and Boudnik}}}
      \newcount\mainNumberOfReferences \mainNumberOfReferences=0
      \def\bibcite#1#2{\global\advance\mainNumberOfReferences by 1\oldbibcite{#1}{#2}}
      \IfFileExists{paper-seb-main.aux}{%
      \input paper-seb-main.aux
      }{\typeout{You will need to run Latex on paper-seb-main.tex because we need its aux file}}
      }
\makeatother

@@

\noindent\vbox{% The table of contents
\def\contents#1#2{\vskip 0mm \noindent{\bfseries #1} \dotfill\ #2\\}
\def\subcontents#1#2{\hbox to 1em{}{#1} \dotfill\ #2\\}
\noindent
\IfFileExists{generated-supplementary.toc}{%
      \input generated-supplementary.toc
      }{\typeout{You will need to run Latex again on paper-seb-supplementary.tex because we need its .toc file}}
}

\newwrite\contentsfile
\immediate\openout\contentsfile=generated-supplementary.toc

@@

\vbox{\noindent\IfFileExists{generated-minicontents.tex}{%
      \input generated-minicontents.tex
      }{\typeout{You will need to run Latex again on paper-seb-supplementary.tex because we need its generated-minicontents.tex file}}}

@@

\immediate\closeout\infofile
\immediate\closeout\contentsfile
\immediate\closeout\minicontentsfile

\end{document}
